---
title: "kmeans"
output: html_document
---

```{r setup, include=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(caret, data.table, ggplot2, ranger, magrittr, dplyr, BBmisc)
```

## load the data for a few seconds


```{r}
columns <- read.table(
    "../data/kddcup.names",
    sep=":",
    skip=1,  # the first column name are the labels, but those are at the end!
    as.is=T
)
column_names <- c(columns[,1], 'label')

connection_events <- read.csv(
    "../data/kddcup.data_10_percent.gz",
    col.names=column_names
)

type <- read.table(
  "http://kdd.ics.uci.edu/databases/kddcup99/training_attack_types",
  sep=" ",
  skip=1
)
fourt = factor(type[[2]])
type[[2]]=="r2l"

setDT(connection_events)  # convert from data.frame to data.table
```


## This code changes strings into number and normalizes the whole data for kmeans classification.


```{r}
connection_events$protocol_type <- as.integer(factor(connection_events$protocol_type))
connection_events$service <- as.integer(factor(connection_events$service))
connection_events$flag <- as.integer(factor(connection_events$flag))
kmeansdata <- connection_events[,1:41]
kmeansdata <- normalize(kmeansdata, method = "standardize", range = c(0, 1), margin = 1L, on.constant = "quiet")
```

## K-means: 5 catogries(Trying to devided the attacks type into 4 automatically)

```{r}
set.seed(20)
Cluster <- kmeans(kmeansdata, 5, nstart = 20)
```


## Add new features or new columns namely is_kmeans_1, is_kmeans_2, ..., is_kmeans_5 and creating new variables named kmeans_1, kmeans_2, ..., kmeans_5

```{r}

for(i in 1:5){
  name = paste0("is_kmeans",i)
  connection_events[[name]]<- as.integer(Cluster$`cluster`==i)
}

for(i in 1:5){
  variablename = paste("kmeans", i, sep = "_")
  columnname = paste("is_kmeans", i, sep = "_")
  assign(variablename, connection_events[connection_events$columnname==1])
}

```

## K-means: 20 catogries (to find some features are new which may be really good for finging features inportance)

```{r}
set.seed(20)
Cluster1 <- kmeans(kmeansdata, 20, nstart = 20)
```

## Add new features

```{r}
model_20 <- connection_events[,c(1:42)]
for(i in 1:20){
  name = paste0("is_kmeans_20_",i)
  model_20[[name]]<- as.integer(Cluster1$`cluster`==i)
}
saveRDS(summary_matrix, file='../data/Kmeans-20-features.rds')

```

# K-Fold Validation

```{r}
model20 <- model_20[,c(42:62)]
k_folds <- createFolds(model_20$label, k=10)
```

This generates 10 sets of indices on the data. These are arranged such that similar amounts of each label are in each set.

To use them for k-fold cross-validation:
  1. Pick each fold one at a time.
  2. Treat this as the indices of the testing set.
  3. Select all other connection events as the training set.
  4. Train your model and get your predictions from the test set.

Example below with random forests.


# Random Forests

Training for all 10 folds:
```{r}
confusion_matrices <-lapply(k_folds, function(fold) {  # approx. 10 minutes
    training_indices <- -fold
    testing_indices <- fold

    training_data <- connection_events[training_indices]
    testing_data <- connection_events[testing_indices]

    model <- ranger(
        label~.,
        data=training_data,
        ## automatically selected since the 'label' column is a factor, but
        ## leave this here for clarity:
        classification=TRUE
    )

    predictions <- predict(model, data=testing_data)

    confusion_matrix <- table(testing_data$label, predictions$predictions)

    return(confusion_matrix)
})
```

```{r}
confusion_matrices
```


```{r}
summary_matrix = Reduce('+', confusion_matrices)
summary_matrix
```

```{r}
saveRDS(summary_matrix, file='../data/20-kmeans-summary-matrix-all-features.rds')

```

```{r}


# Define a function for this to make sure all of our confusion matrices use

# consistent labelling and axis' for true/predicted classes.

make_confusion_matrix <- function(true_values, predicted_values) {

    table(true_values, predicted_values)

}



# Define a function to plot a confusion matrix with a logarithm axis (well,

# log(x+1) to ensure 0 maps to 0). Expects the format output by the

# make_confusion_matrix function above.

plot_confusion_matrix <- function(confusion_matrix) {

    breaks <- c(0, 1, 10, 100, 1000, 10000) * max(confusion_matrix) / 10000

    labels <- sapply(breaks, function(break_value) sprintf("%.0f", break_value))



    confusion_matrix <- as.data.table(confusion_matrix)

    colnames(confusion_matrix) <- c('Predicted', 'True', 'Frequency')



    plot <- ggplot(

            data=confusion_matrix,

            aes(x=Predicted, y=True, fill=Frequency),

            limits=c(0, max(confusion_matrix))

    ) +

        geom_raster() +

        theme(axis.text.x=element_text(angle=90, hjust=1)) +

        scale_fill_gradient(name="Frequency", trans = "log1p", breaks=breaks, labels=labels)



   return(plot)

}



# Example:

plot_confusion_matrix(summary_matrix)
```


